---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: html_document
---

## Assignment 3 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia, focusing on pitch.
In the course of this assignment we will use them to try to automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#packages
## LOAD IN PACKAGES

library(pacman)
p_load(tidyverse, 
       knitr, #kable()
       lmerTest, #lmer()
       caret, #sensitivity() etc.
       MASS, #stepAIC()
       pROC, #ci()
       boot, #inv.logit()
       groupdata2, #fold()
       buildmer, #black-box
       MuMIn)  #dredge()

select <- dplyr::select

## LOAD IN DATA

data <- read.csv("data.csv") %>%
  mutate(ID = factor(ID),
         study = factor(study),
         uniqueness = paste0(ID, diagnosis),
         diagnosis = factor(diagnosis))

clinic <- read.delim("A3_Clinic_Info.txt") %>%
  rename(ID = Subject, 
         study = Study) %>%
  mutate(ID = factor(ID),
         study = factor(study),
         diagnosis = ifelse(Diagnosis == "Control", 0, 1),
         uniqueness = paste0(ID, diagnosis)) %>%
  dplyr::select(ID, study, Gender, uniqueness)


## MERGE
# left join = keeping everyone who appears in "data". Removing those who are in "clinic" only.
# see https://docs.google.com/document/d/1csAffjmezQZHDmsOKzjSiGkD03BWxdDiJLSVOTRC98o/edit
Bear <- inner_join(data, clinic, by = c("uniqueness")) %>%
  dplyr::select(-ID.y, -study.y) %>%
  rename(ID = ID.x, study = study.x) %>%
  mutate(uniqueness = as.numeric(uniqueness))


## CLEAN UP AFTER YOURSELVES
rm(data)
rm(clinic)


## s33d
set.seed(1337)
```

### Question 1: Can you diagnose schizophrenia from pitch range only? If so, how well?

Build a logistic regression to see whether you can diagnose schizophrenia from pitch range only.

Calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve) on a logistic regression using the full dataset. Don't forget the random effects!

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures.

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?

```{r scaling}
# new df not to overwrite
Bear_scaled <- Bear

# MINMAX
for (i in c(6:13)) {
  minc = min(Bear_scaled[,i])
  maxc = max(Bear_scaled[,i])
  Bear_scaled[,i] = (Bear_scaled[,i]-minc)/(maxc-minc)
}

# keep it clean
rm(i, maxc, minc)
```


# Q1: range as predictor
Training model.
Following range model was the most reliable as shown by cross-validation
```{r range model}
## range model
model_range <- glmer(diagnosis ~ range + Gender + (1|ID) + (1|study), 
                     Bear_scaled, family= "binomial")


## PREDICT
# predicting response 
range_pred <- inv.logit(predict(model_range, Bear_scaled, allow.new.levels = TRUE))
range_pred_num <- inv.logit(predict(model_range, Bear_scaled, allow.new.levels = TRUE))

# turning it into 2 level factor
range_pred[range_pred < 0.5] = "0"
range_pred[range_pred >= 0.5] = "1"
```

## ANSWER 1
tidyer output later on (id_uniq_perf)
```{r range performance}
## SENSITIVITY SPECIFICITY
caret::sensitivity(factor(range_pred), factor(Bear_scaled$diagnosis), positive = "1")
caret::specificity(factor(range_pred), factor(Bear_scaled$diagnosis), negative = "0")

## PPV / NPV
PPV <- posPredValue(factor(range_pred), factor(Bear_scaled$diagnosis), positive = "1")
NPV <- negPredValue(factor(range_pred), factor(Bear_scaled$diagnosis), negative = "0")

## ACCURACY
# table to asses
table_range <- table(Bear_scaled$diagnosis, range_pred)

# accuracy function
acc_fun <- function(table) {
  acc = (table[1]+table[4])/sum(table)
  return(round(acc, 2))
}

# accuracy score
acc_fun(table_range)


#ROC curve 
rocCurve <- roc(response = Bear_scaled$diagnosis, 
                predictor = range_pred_num)
pROC::auc(rocCurve)
ci(rocCurve)
plot(rocCurve, legacy.axes = TRUE)

```


# CROSSVALIDATION
-creating folds
```{r folds}
set.seed(1337)
# adds a new column with fold number
Bear_scaled <- fold(Bear_scaled, k = 5, 
             cat_col = 'diagnosis', #balances ratio between shizo/control
             id_col = 'ID') #keeps same people in same folds

```

-define funciton
```{r loop}
# heavily inspired by Ludvig Renbo Olsen
# https://cran.r-project.org/web/packages/groupdata2/vignettes/cross-validation_with_groupdata2.html

crossvalidation_nation <- function(data, k, model_name, dependent, pos, neg) {
  # Initialize empty list for recording performances
  performances_SEN <- c()
  performances_SPE <- c()
  performances_PPV <- c()
  performances_NPV <- c()
  performances_AUC <- c()
  performances_ACC <- c()

  
    # One iteration per fold
  for (fold in 1:k){   #from 1 to k
    
    # Create training set for this iteration
    # Subset all the datapoints where .folds does not match the current fold
    training_set <- data[data$.folds != fold,]
    
    # Create test set for this iteration
    # Subset all the datapoints where .folds matches the current fold
    testing_set <- data[data$.folds == fold,]
    
    ## Train model

    # If there is a random effect,
    # use lmer() to train model
    # else use lm()
    
    model <- glmer(model_name, training_set, family = "binomial")


    ## Test model

    # Predict the dependent variable in the testing_set with the trained model
    predicted <- inv.logit(predict(model, testing_set, allow.new.levels=TRUE))
    
    # Make predicted into factors
    predicted[predicted < 0.5] = "0"
    predicted[predicted >= 0.5] = "1"
    predicted <- factor(predicted)
    
    # Get model performance metrics between the predicted and the observed
    SEN <- caret::sensitivity(predicted, testing_set[[dependent]], positive = pos)
    
    SPE <- caret::specificity(predicted, testing_set[[dependent]], negative = neg)
    
    PPV <- posPredValue(predicted, testing_set[[dependent]], positive = pos)
    
    NPV <- negPredValue(predicted, testing_set[[dependent]], negative = neg)
    
    predicted <- as.numeric(predicted)
    rocCurve <- roc(response = testing_set$diagnosis, predictor = predicted)
    AUC <- pROC::auc(rocCurve)
    
    testing_set$predicted <- predicted
    testing_correct <- filter(testing_set, diagnosis == predicted)
    ACC <- nrow(testing_correct) / nrow(testing_set)
    
    # Add the to the performance list
    performances_SEN[fold] <- SEN
    performances_SPE[fold] <- SPE
    performances_PPV[fold] <- PPV
    performances_NPV[fold] <- NPV
    performances_AUC[fold] <- AUC
    performances_ACC[fold] <- ACC

  }

  # Return the mean of the recorded RMSEs
  return(cbind.data.frame('model' = model_name,
           'SEN' = mean(performances_SEN, na.rm = T),
           'SPE' = mean(performances_SPE, na.rm = T),
           'PPV' = mean(performances_PPV, na.rm = T),
           'NPV' = mean(performances_NPV, na.rm = T),
           'AUC' = mean(performances_AUC, na.rm = T),
           'ACC' = mean(performances_ACC, na.rm = T)))

}
```


## ID vs uniqueness debate
Reasoning behind diagnosis ~ range + Gender + (1|ID) + (1|study)

Q: Gender as a fixed effect?
A: yes, it shows a more reliable performance than the model which excludes gender alltogether.

Decrease in SEN is acompanied by increase in SPE and overall accuracy is also slighlty higher. Similar tendency happens between PPV and NPV. 

The model with gender as random effect is performing similarly good. The reason for it's rejection is based on a marginal difference in SEN and SPE (gender as fixed effect is performing slightly better) and for conceptual reasons (gender is an obvious voice predictor)

```{r}
set.seed(1337)

#Q1 ID
rm1 <- 'diagnosis ~ range + (1|ID) + (1|study)'
rm2 <- 'diagnosis ~ range + Gender + (1|ID) + (1|study)'
rm3 <- 'diagnosis ~ range + (1|Gender) + (1|ID) + (1|study)'

id_uniq_list <- list(rm1, rm2, rm3)

id_uniq_perf <- map_df(
            #list to loop trough
            id_uniq_list,
            #function to use
            crossvalidation_nation,
            # arguments
            data = Bear_scaled, 
            k = 5,
            dependent = "diagnosis", 
            pos = "1", 
            neg = "0")
```


### Question 2 - Which single acoustic predictor is the best predictor of diagnosis?
# model list

```{r model list}
# creating model calls and putting them into list for later use
m1 <- "diagnosis ~ range + Gender + (1|ID) + (1|study)"
m2 <- "diagnosis ~ sd + Gender + (1|ID) + (1|study)"
m3 <- "diagnosis ~ mean + Gender + (1|ID) + (1|study)"
m4 <- "diagnosis ~ median + Gender + (1|ID) + (1|study)"
m5 <- "diagnosis ~ iqr + Gender + (1|ID) + (1|study)"
m6 <- "diagnosis ~ mean_abs + Gender + (1|ID) + (1|study)"
m7 <- "diagnosis ~ coef_var + Gender + (1|ID) + (1|study)"

model_list <- list(m1, m2, m3, m4, m5, m6, m7)
```


### cross-validation loop

```{r output}
set.seed(1337)
cross_table <- map_df(
            #list to loop trough
            model_list,
            #function to use
            crossvalidation_nation,
            # arguments
            data = Bear_scaled, 
            k = 5,
            dependent = "diagnosis", 
            pos = "1", 
            neg = "0")

```

### best predictor
Median seems to be the best predictor. It scores the highest on every performance meassure except for accuracy.
```{r}
mean_per <- cross_table %>%
  select(-AUC) %>%
  rowwise() %>%
  mutate(mean_per = mean(c(SEN, SPE, PPV, NPV)),
         sd_per = sd(c(SEN, SPE, PPV, NPV)))

```


### Question 3 - Which combination of acoustic predictors is best for diagnosing schizophrenia?
this section contains:
- intuition approach (try to guess which model would be nice)
- black box approach (automated model selection)
- not working: brute force approach (p-hacking loop)

this section does not contai (but should):
- NO OUT-OF-SAMPLE ERROR

## intuition approach
Models were build based on performance of individual predictors:
median > mean
iqr > range

However, both models perform very similarly and that is: not good enough. 
```{r}
## models

# model based on momentarily intuitions
intuition <- glmer(diagnosis ~ median + iqr + coef_var + Gender +
                     (1|ID) + (1|study), Bear_scaled, family = "binomial")

# model based on what we though would fuck up
bad <- glmer(diagnosis ~ mean + range + coef_var + Gender + 
                     (1|ID) + (1|study), Bear_scaled, family = "binomial")



## cross-validation

intuition_cross = crossvalidation_nation(data = Bear_scaled, k = 5, 
                       model_name = "diagnosis ~ median + iqr + coef_var + Gender +
                     (1|ID) + (1|study)", 
                       dependent = "diagnosis", pos = "1", neg = "0")

bad_cross = crossvalidation_nation(data = Bear_scaled, k = 5, 
                       model_name = "diagnosis ~ mean + range + coef_var + Gender + 
                     (1|ID) + (1|study)", 
                       dependent = "diagnosis", pos = "1", neg = "0")

```



## BLACK BOX APPROACH
Using automated model selection to generate the best formula (based on AIC)
-order, backward
```{r}
a2 <- buildmer(diagnosis ~ mean * sd * range * iqr * 
                median * mean_abs * coef_var * se * 
                Gender + (1|ID) + (1|study) + 
                (mean|ID) + (mean|study) +
                (sd|ID) + (sd|study) +
                (iqr|ID) + (iqr|study) +
                (median|ID) + (median|study) +
                (mean_abs|ID) + (mean_abs|study) +
                (coef_var|ID) + (coef_var|study) +
                (se|ID) + (se|study), 
              Bear_scaled, family = "binomial")


buildmer_best <- "diagnosis ~ se + mean + iqr + mean_abs + coef_var + Gender + (1 | study)"

best_model <- glmer(diagnosis ~ se + mean + iqr + mean_abs + coef_var + Gender + (1 | study),
                    Bear_scaled, family = "binomial")

best <- crossvalidation_nation(data = Bear_scaled, k = 5, 
                               model_name = buildmer_best, 
                               dependent = "diagnosis", pos = "1", neg = "0")
```


### visualizing black box
```{r}
Bear_best <- Bear_scaled %>%
  ungroup() %>%
  mutate(pred_best_num = inv.logit(predict(best_model, Bear_scaled, 
                               allow.new.levels=TRUE)),
         pred_best = ifelse(pred_best_num < 0.5, 0, 1),
         pred_tab = case_when(pred_best == 0 & diagnosis == 0 ~ "true negative",
                              pred_best == 0 & diagnosis == 1 ~ "false negative",
                              pred_best == 1 & diagnosis == 0 ~ "false positive",
                              pred_best == 1 & diagnosis == 1 ~ "true positive"))

labels <- as.data.frame(table(Bear_best$pred_tab)) %>%
  rename(pred_tab = Var1)

```


#### confusion matrix plotted 
```{r}
Bear_best %>%
  mutate(diagnosis = ifelse(diagnosis == 0, "control", "schizophrenic"),
         pred_best = ifelse(pred_best == 0, "control", "schizophrenic")) %>%
  ggplot(aes(diagnosis, pred_best_num, fill = factor(pred_tab))) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(panel.grid.major.x = element_blank()) +
  labs(title = "Model performance", 
       subtitle = "diagnosis ~ se + mean + iqr + mean_abs + coef_var + Gender + (1 | study)",
       fill = "model prediction", 
       y = "number of model predictions", 
       x = "true diagnosis")
```

#### more plots
```{r}
Bear_best %>%
  ggplot(aes(diagnosis, fill = factor(pred_best))) +
  geom_bar(position = "fill") +
  theme_bw() +
  labs(title = "Model performance", 
       subtitle = "diagnosis ~ se + mean + iqr + mean_abs + coef_var + Gender + (1 | study)",
       fill = "model prediction", 
       x = "true diagnosis")
```


## BRUTE FORCE
BAD SECTION, PLESE NO
```{r phack brute}
fix_bear <- names(Bear_scaled)[6:13] %>%
  paste(collapse=" + ") %>%
  paste0("diagnosis ~ ", .)
```

```{r phack manual and lame}
g1 <- "diagnosis ~ median + Gender + (1|ID) + (1|study)"
g2 <- "diagnosis ~ median + Gender + (median|ID) + (1|study)"
g3 <- "diagnosis ~ median + Gender + (1|ID) + (median|study)"
g4 <- "diagnosis ~ median + Gender + (median|ID) + (median|study)"

g5 <- "diagnosis ~ se + mean + iqr + mean_abs + coef_var + Gender + (1|ID) + (1|study)"
g6 <- "diagnosis ~ se + mean + iqr + mean_abs + coef_var + Gender + (mean|ID) + (1|study)"
g7 <- "diagnosis ~ se + mean + iqr + mean_abs + coef_var + Gender + (1|ID) + (mean|study)"
g8 <- "diagnosis ~ se + mean + iqr + mean_abs + coef_var + Gender + (mean|ID) + (median|study)"
g9 <- "diagnosis ~ se + mean + iqr + mean_abs + coef_var + Gender + (mean|study)"






gmodel <- list(g1, g2, g3, g4)
gmodel2 <- list(g5, g6, g7, g8, g9)
```

```{r}
set.seed(1337)
g_table <- map_df(
            #list to loop trough
            gmodel2,
            #function to use
            crossvalidation_nation,
            # arguments
            data = Bear_scaled, 
            k = 5,
            dependent = "diagnosis", 
            pos = "1", 
            neg = "0")

mean_g2 <- g_table %>%
  select(-AUC) %>%
  rowwise() %>%
  mutate(mean_per = mean(c(SEN, SPE, PPV, NPV)),
         sd_per = sd(c(SEN, SPE, PPV, NPV)))
```


### Question 4: Properly report the results

METHODS SECTION: how did you analyse the data? That is, how did you extract the data, designed the models and compared their performance?

RESULTS SECTION: can you diagnose schizophrenia based on voice? which features are used? Comment on the difference between the different performance measures.

### Bonus question 5

You have some additional bonus data involving speech rate, pauses, etc. Include them in your analysis. Do they improve classification?

### Bonus question 6

Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them.
